Postural changes mediate children's visual access to social information

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
both detectors -- tables (need formatting), some edits to figure captions

```{r xtable, results="asis"}
tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")

@@

```{r xtable, results="asis"}
 tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3,
                       caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.")

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
\end{table}

@@

```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ("Look, a Zem"; +/- 2 seconds around each utternace) as a function of infants's posture. Each point represents a naming instance for each child in the dataset."  }
ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
  geom_boxplot() + 
  ylab("Prop. face detections") + 
  xlab("Posture") + 
  theme_bw() 
```
## Access to Faces During Labeling Events
Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.  Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections. We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.

>>>>

```{r MTCNNModel, results="asis"}
tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")

@@

```{r OPModel, results="asis"}
 tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3,
                       caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.")

@@

<!-- \begin{table}[H] -->
<!-- \centering -->
<!-- \begin{tabular}{rrrrr} -->
<!--   \hline -->
<!--  & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\  -->
<!--   \hline -->
<!-- (Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\  -->
<!--   Age & 0.170 & 0.061 & 2.766 & 0.006 \\  -->
<!--   Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\  -->
<!--   Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\  -->
<!--   Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\  -->
<!--   Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\  -->
<!--   Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\  -->
<!--    \hline -->
<!-- \end{tabular} -->
<!-- \caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).}  -->
<!-- \end{table} -->

@@

```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture. Each point represents a naming instance for each child in the dataset."}
ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
  geom_boxplot() + 
  ylab("Prop. face detections") + 
  xlab("Posture") + 
  theme_bw() 
```
## Access to Faces During Labeling Events
Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.  Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections. We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
new paragraph
both detectors -- tables (need formatting), some edits to figure captions

\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

@@

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
state-of-the-art computer vision algorithms to detect the proportion of
faces in infants' environments. We find that infants' postural changes
largely mediate infants' access to faces, hinting that these postural
and locomotive developments facilitate infants' emerging linguistic and
social capacities. Broadly, we suggest that the combined use of
head-mounted cameras and the application of novel deep learning
algorithms is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

@@

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
also impact how children interact with their mothers; walking
(vs.~crawling) infants make different kinds of object-related bids for
attention from their mothers and tend to hear more action directed
statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
2014). Further, in an observational study, Walle \& Campos (2014) found
that children who were able to walk had both higher receptive and
productive vocabularies. On their account, children's ability to stand
and independently locomote may fundamentally change their ability to
access social information (e.g., faces, gaze) relative to children who
are still crawling and sitting. In other words, the ability to access
more detailed social information may allow infants to learn words
quicker and more efficiently, facilitating language growth.

Recent work has begun to use egocentric, head-mounted cameras to
document the visual experiences of infants and children --- which even
for walking children are radically different from the adult perspective
(and not easily predicted by our own adult intuitions) (Clerkin, Hart,
Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith,
2008). Children's views tend to be restricted and to be dominated by
objects and hands much more than that of adults (Yoshida \& Smith,
2008), and both computational and empirical work suggest that this
restricted viewpoint may be more effective for learning objects and
their labels than the comparable adult perspective (Bambach, Crandall,
Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in press). Further,
recent work also suggests dramatic changes in the child's perspective
over the first two years of life, as views transition from primarily
containing a close up view of faces to capturing views of hands paired
with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).

Here, we directly ask whether the postural and locomotive developments
change the availability of the social information relevant for word
learning. To do so, we recorded the visual experience of a group of
infants and children using head-mounted cameras during a brief
laboratory free-play session; children's posture and orientation
relative to their caregiver were also recorded from a third-person
perspective and hand-annotated. We capitalize on recent improvements in
face detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces in the
child's visual environment, both overall and relative to naming events
by their caregivers.

@@

We created a set of custom annotations that described the child's
physical posture (e.g.~standing) and the orientation of the caregiver
relative to the child (e.g.~far away). The child's posture was
categorized as being held/carried, prone (crawling or lying), sitting,
or standing. The caregiver's orientation was characterized as being
close to the child, far from the child, and a global category of
caregiver behind the child. For the first two annotations (close/far
from the child), the caregiver could either be to the the front or to
the side of the child. All annotations were made using OpenSHAPA/Datavyu
software (K. Adolph, Gilmore, Freeman, Sanderson, \& Millman, 2012), and
times when the child was out of view of the tripod camera were marked as
uncodable and were excluded from these annotations.

@@

Measuring the availability of caregiver's faces across development was
an important goal of this study. We used a total of three face detection
systems to explore infants' changing access to social information and to
avoid the cost of hand-annotating every frame. We first measured the
performance of the most commonly-used and widely available face
detection algorithms (Viola-Jones). We used this as a benchmark for
performance, and while it can achieve impressive accuracy in some
situations, it is notoriously bad at dealing with occluded faces
(Scheirer, Anthony, Nakayama, \& Cox (2014)). We next capitalized on
recent improvements in computer vision, testing the performance of two
state-of-the-art face detectors that both made use of Convolutional
Neural Networks (CNNs) to extract face information.

@@

The first face detection system made use of a series of Haar
feature-based cascade classifiers (Viola-Jones, Viola \& Jones (2004))
applied to each individual frame. This detector provided information
about the presence of a face as well as its size and position.

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc.) (Cao et al.
(2017), Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna,
Kanade, \& Sheikh (2016)) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose, as any half of a face necessarily contains a
nose.\\

@@

In order to evaluate the performance of these detectors, we constructed
a ``gold set'' of frames by hand labeling both a sample of frames with a
high density of faces (as reported by the detectors) and a random sample
from the remaining frames. This was done so as to not bias our
evaluation by the relatively rare appearance of faces in the dataset. A
face was present in a frame if at least half of the face was showing.
Precision (hits / hits + false alarms), recall (hits / hits + misses),
and F-score (harmonic mean of previous measures) were calculated and are
reported in Table 2. Both OpenPose and MTCNN detectors performed
relatively well on the gold set, with MTCNN outperforming OpenPose on
the random sample but trailing behind in the high density sample. Due to
the high performance of these two detectors we have included the results
of both in the analyses to follow. The ViolaJones detector did not
perform as well on either gold set, so the results associated with this
detector were not included.

@@

\subsection{Model}\label{model}

To formalize these observations, we fit a generalized logistic
mixed-effect model to all with the presence/absence of a face on every
frame as the dependent variable, and participant's age, orientation, and
posture as predictors. Interactions between predictors were not included
as this maximal model failed to converge. A summary of the coefficients
of this model can be found in Table 3. While age remained a significant
predictor even when accounting for the effects of infants' posture and
caregivers' orientation, it held relatively less predictive power.
Overall, these results confirm that infants' access to faces is heavily
infleunced by their own posture and their caregiver's orientation
towards them.

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
\end{table}

@@

\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}

>>>>

\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Ally Kraus} \\ \texttt{XX@XX} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

@@

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
state-of-the-art computer vision algorithms to detect the proportion of
faces in infants' environments. We find that infants' posture and
orientation to their caregiver largely mediate infants' access to faces,
suggesting that these postural and locomotive developments facilitate
infants' emerging linguistic and social capacities. Broadly, we suggest
that the combined use of head-mounted cameras and the application of
novel deep learning algorithms is a promising avenue for understanding
the statistics of infants' visual and linguistic experience.

@@

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
also impact how children interact with their mothers; walking
(vs.~crawling) infants make different kinds of object-related bids for
attention from their mothers and tend to hear more action directed
statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
2014). Further, in an observational study, Walle \& Campos (2014) found
that children who were able to walk had both higher receptive and
productive vocabularies. On their account, children's ability to stand
and independently locomote may fundamentally change their ability to
access social information (e.g., faces, gaze) relative to children who
are still crawling and sitting. In other words, the ability to walk and
move around independently and subsequently access more detailed social
information may allow infants to learn words quicker and more
efficiently, facilitating language growth.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
document visual experiences of infants and children --- which even for
walking children are radically different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be restricted and to be dominated by objects
and hands much more than that of adults (Yoshida \& Smith, 2008), and
both computational and empirical work suggest that this restricted
viewpoint may be more effective for learning objects and their labels
than the comparable adult perspective (Bambach, Crandall, Smith, \& Yu,
2017; D. Yurovsky, Smith, \& Yu, in press). Further, recent work also
suggests dramatic changes in the child's perspective over the first two
years of life, as views transition from primarily containing a close up
view of faces to capturing views of hands paired with the objects they
are acting on (Fausey, Jayaraman, \& Smith, 2016).

Here, we directly ask whether postural and locomotive developments
change the availability of the social information relevant for word
learning. To do so, we recorded the visual experience of a group of
infants and children using head-mounted cameras during a brief
laboratory free-play session; children's posture and orientation
relative to their caregiver were also recorded from a third-person
perspective and hand-annotated. We capitalize on recent improvements in
face detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces in the
child's visual environment, both overall and relative to naming events
by their caregivers.

@@

We created a set of custom annotations that described the child's
physical posture (e.g.~standing) and the orientation of the caregiver
relative to the child (e.g.~far away). The child's posture was
categorized as being held/carried, prone (crawling or lying), sitting,
or standing. The caregiver's orientation was characterized as being
close to the child, far from the child, and a global category of
caregiver behind the child. For the first two annotations (close/far
from the child), the caregiver could either be to the the front or to
the side of the child. All annotations were made by a trained coder
using the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman,
Sanderson, \& Millman, 2012), and times when the child was out of view
of the tripod camera were marked as uncodable and were excluded from
these annotations.

@@

We used a total of three face detection systems to explore infants'
changing access to social information and to avoid the cost of
hand-annotating every frame. We first measured the performance of the
most commonly-used and widely available face detection algorithms
(Viola-Jones). We used this as a benchmark for performance, and while it
can achieve impressive accuracy in some situations, it is notoriously
bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox,
2014). We next capitalized on recent improvements in computer vision,
testing the performance of two state-of-the-art face detectors that both
made use of Convolutional Neural Networks (CNNs) to extract face
information. The first algorithm was specifically optimized for face
detection, and the second algorithm was optimized to extract information
about the position of agent's bodyparts.

@@

The first face detection system made use of a series of Haar
feature-based cascade classifiers (Viola-Jones, (Viola \& Jones, 2004))
applied to each individual frame. This detector provided information
about the presence of a face as well as its size and position.

@@

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose, as any half of a face necessarily contains a
nose.\\

@@

In order to evaluate the performance of these detectors, we constructed
a ``gold set'' of frames by hand labeling both a sample of frames with a
high density of faces (as reported by the detectors) and a random sample
from the remaining frames. This was done so as to not bias our
evaluation by the relatively rare appearance of faces in the dataset. A
face was present in a frame if at least half of the face was showing.
Precision (hits / hits + false alarms), recall (hits / hits + misses),
and F-score (harmonic mean of previous measures) were calculated and are
reported in Table 2.

Both OpenPose and MTCNN detectors performed relatively well on the gold
set, with MTCNN outperforming OpenPose on the random sample but trailing
behind in the high density sample. Due to the high performance of these
two detectors we have included the results of both in the analyses to
follow. The ViolaJones detector did not perform well on either gold set,
so the results from this detector were not included.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
  \hline
 & Algorithm & Sample\ Type & P & R & F \\ 
  \hline
1 & MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
  2 & MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
  3 & OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
  4 & OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
  5 & ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
  6 & ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
   \hline
\end{tabular}
\caption{Detector performance.} 
\end{table}

@@

To formalize these observations, we fit a generalized logistic
mixed-effect model to all with the presence/absence of a face on every
frame as the dependent variable, and participant's age, orientation, and
posture as predictors. Interactions between predictors were not included
as this maximal model failed to converge. A summary of the coefficients
of this model can be found in Table 3 for the MTCNN detections, though
we found the same pattern of results for the OpenPose detections. While
age remained a significant predictor even when accounting for the
effects of infants' posture and orientation, it accounted for the least
amount of variance. Overall, these results suggest that, instead,
infants' access to faces is heavily infleunced by their own posture and
theirorientation towards their caregiver.

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  scale(age.at.test) & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  postureprone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  posturesit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  posturestand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  orientationclose & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  orientationfar & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
  scale(age.at.test) & 0.170 & 0.061 & 2.766 & 0.006 \\ 
  postureprone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
  posturesit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
  posturestand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
  orientationclose & 1.549 & 0.021 & 72.757 & 0.000 \\ 
  orientationfar & 2.258 & 0.022 & 101.138 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.} 
\end{table}

@@

\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture. Each point represents a naming instance for each child in the dataset.}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}
